---
title: "DATA 605 HW4 YINA QIAO"
output: html_document
---
Requirement: With the attached data file, build and visualize  eigenimagery that accounts for 80% of the variability.  Provide full R code and discussion.

## File Loading and Preprocessing

Provied data contains 17 images. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
memory.size(max = TRUE)
library(doParallel)
library(foreach)
library(jpeg)
library(EBImage)
```

```{r}
# Define the file path where the images are located
files <- list.files(path = "/Users/linda/Desktop/elearning/CUNY-MSDS/data 605/w4/jpg", pattern = "\\.jpg")
```

## Image Resizing and Plotting:


```{r}
# Parameters for image processing
height <- 1200
width <- 2500
scale <- 20

# Function to plot a JPEG image
plot_jpeg <- function(path, add = FALSE) {
  require('jpeg')
  jpg <- readJPEG(path, native = TRUE)
  res <- dim(jpg)[2:1]
  if (!add) {
    plot(1, 1, xlim = c(1, res[1]), ylim = c(1, res[2]),
         asp = 1, type = 'n', xaxs = 'i', yaxs = 'i',
         xaxt = 'n', yaxt = 'n', xlab = '', ylab = '',
         bty = 'n')
  }
  rasterImage(jpg, 1, 1, res[1], res[2])
}

# Initialize an array to store the images
im <- array(rep(0, length(files) * height/scale * width/scale * 3),
            dim = c(length(files), height/scale, width/scale, 3))

# Load and resize images
for (i in seq_along(files)) {
  tmp <- paste0("/Users/linda/Desktop/elearning/CUNY-MSDS/data 605/w4/jpg/", files[i])
  temp <- EBImage::resize(readJPEG(tmp), height/scale, width/scale)
  im[i,,,] <- array(temp, dim = c(1, height/scale, width/scale, 3))
}


# Set up plotting parameters
par(mfrow = c(3, 3))
par(mai = c(0.3, 0.3, 0.3, 0.3))

# Plot the first 9 images only
for (i in 1:9) {
  plot_jpeg(writeJPEG(im[i,,,]))
}

```

## Principal Component Analysis (PCA):



```{r}
# Prepare the data for PCA
dim(im) <- c(length(files), height * width * 3 / scale^2)
mypca <- princomp(t(as.matrix(im)), scores = TRUE, cor = TRUE)

# Calculate and print the variance
cat("Variance explained by each component:\n")
print(mypca$sdev^2 / sum(mypca$sdev^2))

# Calculate the cumulative variance
cumulative_variance <- cumsum(mypca$sdev^2 / sum(mypca$sdev^2))
cat("Cumulative variance explained:\n")
print(cumulative_variance)

# Find the number of components explaining 80% variance
num_components_80 <- which(cumulative_variance >= 0.80)[1]
cat("Number of components explaining at least 80% variance: ", num_components_80, "\n")

```

## Visualization of Eigenimages explaining 80% variance:



```{r}
# Reshape the PCA scores to visualize the eigenimages
mypca2 <- t(mypca$scores)
dim(mypca2) <- c(length(files), height/scale, width/scale, 3)

# Set up plotting parameters for eigenimages
par(mfrow = c(1, num_components_80))
par(mai = c(0.001, 0.001, 0.001, 0.001))

# Plot the eigenimages explaining 80% variance
for (i in 1:num_components_80) {
  plot_jpeg(writeJPEG(mypca2[i,,,], quality = 1, bg = "white"))
}

```

### Discussion
**The analysis finds that only three components are needed to explain at least 80% of the variance in the image dataset. This suggests that a substantial portion of the variability in the images can be represented by a relatively small number of components. **
