
---
title: " DATA 605 HW10"
author: "Yina Qiao"

output: html_document
---

## Problem Statement

Smith is in jail with 1 dollar and can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability 0.4 and loses A dollars with probability 0.6.


### (a) Timid Strategy

Smith bets 1 dollar each time. The probability of reaching 8 dollars before losing all his money 


- The example is well-suited to apply the well-known gambler ruin's problem
- where a player, starting with a certain stake, aims to reach a goal before depleting their resources, under the conditions of winning and losing fixed amounts each round.



```{r}
z <- 1  # Smith's starting amount
M <- 8  # Target amount to win
q <- 0.6  # Probability of losing a bet
p <- 0.4  # Probability of winning a bet

if (p != q) {
  qz <- ((q/p)^z - 1) / ((q/p)^M - 1)
} else {
  qz <- z / M
}

qz
```

### simulation 
```{r}
start_dollars <- 1
out_jail <- c()

for (j in 1:100000) {
  samples <- c(start_dollars)
  for (x in 1:100) {
    samples <- append(samples, sample(c(1,-1), size=1, replace=T, prob=c(.4,.6)))
    if (sum(samples) >= 8){
      out_jail <- append(out_jail, 1)
      break
    } else if (sum(samples) <= 0) {
      break
    }
  }
}  

sum(out_jail) / 100000
```



### (b) Bold Strategy

Smith bets as much as possible but not more than necessary to reach 8 dollars.



- Direct calculation: \( p_1 = 0.4^3 \), considering only wins needed to reach 8 dollars.


```{r bold_strategy, echo=TRUE}
p1_bold <- 0.4^3
cat("Probability with bold strategy:", p1_bold)
```

### simulation

```{r}
start_dollars <- 1
out_jail <- c()

for (j in 1:100000) {
  samples <- c(start_dollars)
  for (x in 1:100) {
    samples <- append(samples, sample(c(sum(samples),-sum(samples)), size=1, replace=T, prob=c(.4,.6)))
    if (sum(samples) >= 8){
      out_jail <- append(out_jail, 1)
      break
    } else if (sum(samples) <= 0) {
      break
    }
  }
}  

sum(out_jail) / 100000
```

### (c) Comparison of Strategies

The strategy with a higher probability \( p_1 \) gives Smith a better chance of getting out of jail.\n\
Probability with timid strategy : 0.02\n\
Probability with bold strategy: 0.064,\n\
Therefore bold strategy gives a smith a better chance.


